---
title: "ds202_hw5"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1.The code below downloads the data for Ames only and saves a local copy to your machine. The first time you run this code, make sure you have a good internet connection and enough space on your machine (~50 MB). If the local copy exists, re-knitting the file will load the (binary) file from disk and be much faster.
```{r}
library(readr)
if (file.exists("alcohol.rds")) {
  data <- readRDS("alcohol.rds")
} else {
  url <- "https://xdaiisu.github.io/ds202materials/data/ames.csv"
  data <- readr::read_csv(url)
  saveRDS(data, file="alcohol.rds")
}
```

2.Data cleaning:
·extract geographic latitude and longitude from the variable Store Location
·check variable types. Pick five variables that need to be converted to a different type and fix those.
·extract year, month and day from the variable Date
```{r}
library(ggplot2)
library(dplyr)
library(tidyr)

dataClean <- readRDS("alcohol.rds")
dataClean <- suppressWarnings(separate(dataClean, "Store Location", c("Address", "Coordinates"), sep = "\\("))
dataClean <- separate(dataClean, "Coordinates", c("latitude", "longitude"), sep = ",")
dataClean<- separate(dataClean, "longitude", c("longitude", "empty"), sep = "\\)")
dataClean <- dataClean %>% select(-c("empty"))
dataClean$latitude <- as.numeric(dataClean$latitude)
dataClean$longitude <- as.numeric(dataClean$longitude)
dataClean$`Vendor Number` <- as.numeric(dataClean$`Vendor Number`)
dataClean$`Store Name` <- as.factor(dataClean$`Store Name`)
dataClean$`Category Name` <- as.factor(dataClean$`Category Name`)
dataClean <- dataClean %>% mutate("Real Date" = Date) %>% separate("Date", c("Month", "Day", "Year"), sep="/")
```

3.First overview:
·Plot a scatterplot of lat and long of store locations.

```{r}
data <- dataClean
suppressWarnings(ggplot(data, aes(x = latitude, y = longitude)) + geom_point())
```

·Provide a visual breakdown of the liquor category (by Category Name). Include volume sold in the breakdown.

```{r}
visual <- suppressWarnings(data %>% group_by(`Category Name`) %>% summarise(total = (sum(`Volume Sold (Liters)`)/1000)))
visual <- visual[order(-visual$total), ]
visual <- head(visual, 20)
ggplot(visual, aes(x= reorder(`Category Name`, -total), y=total)) + geom_col() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + xlab("liquor category (by Category Name)") + ylab("volume sold in the breakdown") + ggtitle("Visual breakdown of the liquor category (by Category Name)")
```


4.Find the daily sales liquor sales in Ames in 2018. Summarize number of sales, volume of liquor sold and amount of money spent. Plot volume sold by day (use a scatterplot of volume by day and facet by month). Describe any patterns. Find the dates for ISU football home games in Fall 2018. Can you see a pattern?

```{r}
#Summarize number of sales, volume of liquor sold and amount of money spent.
sales2018 <- data %>% filter(Year == 2018)
sum2018 <- sales2018 %>% summarise(numSales = sum(sales2018$`Bottles Sold`), volLiquor = sum(sales2018$`Volume Sold (Gallons)`), amtMoney = sum(sales2018$`Sale (Dollars)`))
sum2018

#Plot volume sold by day (use a scatterplot of volume by day and facet by month).
sales2018$Day <- as.numeric(sales2018$Day)
ggplot(sales2018, aes(x=Day, y=`Volume Sold (Gallons)`)) + geom_point() + facet_wrap(~Month) + scale_x_continuous(breaks=seq(0,30,6)) + xlab("Days of Months") + ylab("Volume sold") + ggtitle("Liquor volume sold by day")
```

Describe any patterns:
There are 3-4 peeks in each graph. We can tell that it should be the weekends people buy liquor.The data seems missing alot in October. 

Find the dates for ISU football home games in Fall 2018:
9/1/18, 9/15/18, 9/22/18, 10/13/18, 10/27/18, 11/10/18, 11/24/18, 12/1/18

The football game mainly held in September to December. But we are missing the data in September and most October. The pattern cannot be told this time.